---
title: "How to use your GPU for things other than gaming"
author:
  - Andrew Chen
output:
  xaringan::moon_reader:
    css: ["AC_theme.css", "AC_theme_add.css"]
    nature:
      highlightStyle: tomorrow-night-bright
      beforeInit: "macros.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
header_includes:
  - \usepackage{amsmath}
  - \usepackage{bm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2, # higher quality plots
  dev = "svg", # vector graphics device
  dev.args=list(bg="transparent") # transparent plot backgrounds
)

library(ggplot2)
# Dark Theme
theme_set(theme_minimal())
theme_update(
  panel.background = element_rect(fill = "transparent", colour = NA),
  plot.background = element_rect(fill = "transparent", colour = NA),
  legend.background = element_rect(fill = "transparent", colour = NA),
  legend.key = element_rect(fill = "transparent", colour = NA),
  text = element_text(colour = "white"),
  axis.text = element_text(colour = "white"))

library(MDMR)
```

## Why do we have GPUs
- Graphical processing units (GPUs) were named by Sony for the PlayStation console
- Intended for huge number of simple parallel operations
    - e.g. increasing the brightness of every pixel on a screen
--
    - or most tasks involving rendering, lighting, etc.


.center[
<video autoplay muted>
    <source src="https://thumbs.gfycat.com/FelineHopefulCassowary-mobile.mp4" type="video/mp4" />
</video>
]
---

## Ray-tracing example
.center[
<video autoplay muted>
    <source src="https://thumbs.gfycat.com/ImaginativeSlimyAlleycat-mobile.mp4" type="video/mp4" />
</video>
]
---

## But there are other uses
- General-purpose computing on GPUs
    - Enables CPU computations on GPUs
    - Popularized recently by frameworks such as OpenCL and Nvidia CUDA
- Useful for calculations in
    - Matrix/vector operations
    - Fast fourier transforms
    - Neural networks
    - Image segmentation/processing<sup>1</sup>

.footnote[
<sup>1</sup>[Smistad et al., 2015](https://www.sciencedirect.com/science/article/pii/S1361841514001819)
]
---

## CPUs vs GPUs
- Nvidia GeForce RTX 2080 Ti has 4,352 GPU cores at a clock speed of 1,350MHz
    - A high-end CPU, say the Intel Core i7-9700K, has 8 cores at a clock speed of 3,600MHz
--
- GPUs good at applying same operation to many data units[<sup>1</sup>](https://graphics.stanford.edu/~mhouston/public_talks/R520-mhouston.pdf)
    - Optimized for parallel processing
--
- CPUs better for executing a smaller number of complex tasks
    - Optimized for serial processing
---

## Considerations for GPU computing
- OpenCL vs Nvidia CUDA
    - CUDA written only for Nvidia GPUs
    - Mac support for CUDA ended[<sup>1</sup>](https://gizmodo.com/apple-and-nvidia-are-over-1840015246)
--
- Single-precision vs double-precision
    - Loss of performance for double-precision computing
--
- Object storage
    - GPUs are generally less capable of storing large objects to memory
    - Moving object from CPU to GPU and back is slow
--
- Computer hardware
    - Low-end GPU may not see performance gains over high-end CPU
    - Available software depends on your system specifications
---

## gpuR: GPU matrix computations in R
- `gpuR` provides tools to perform basic matrix operations using the GPU
- Simply need to specify which device to use
```{r load, include=FALSE}
library(gpuR)
```

```{r}
library(gpuR) # uses OpenCL, works on Mac
listContexts()
setContext(3) # select GPU to use
```

---

## gpuR: Matrix inversion
```{r}
# using base R
p = 2000

system.time({
  A <- matrix(runif(p^2), p, p)
  solve(A)
})[3]

# using gpuR
system.time({
  A <- gpuMatrix(runif(p^2), p, p)
  solve(A)
})[3]
```
---

## gpuR: Small matrix inversion
```{r}
# using base R
p = 200

system.time({
  A <- matrix(runif(p^2), p, p)
  solve(A)
})[3]

# using gpuR
system.time({
  A <- gpuMatrix(runif(p^2), p, p)
  solve(A)
})[3]
```
---

## gpuR: Matrix inversion comparison
```{r, eval=FALSE, echo=FALSE}
inv <- sapply(c(seq(50, 450, 50), seq(500, 2500, 500)), function(p) c(
  p,
  system.time({
    A <- matrix(runif(p^2), p, p)
    solve(A)
  })[3],
  system.time({
    A <- gpuMatrix(runif(p^2), p, p)
    solve(A)
  })[3]
))

save(inv, file = "gpuR_inv.Rdata")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
load("gpuR_inv.Rdata")

ggplot(data.frame(p = rep(inv[1,], 2), Time = log(c(inv[2,], inv[3,])),
                  Method = c(rep("Base", ncol(inv)), rep("GPU", ncol(inv))))) +
  geom_line(aes(x = p, y = Time, group = Method, color = Method)) + 
  labs(y = "log2(Time)")
  scale_y_continuous("log2")
```
---

## gpuR: Linear models
```{r}
# using base R
n = 2000
p = 1000

system.time({
  X <- matrix(runif(n*p), n, p)
  beta <- matrix(rnorm(p), p, 1)
  y <- X %*% beta + matrix(rnorm(n), n, 1)
  solve(t(X) %*% X) %*% t(X) %*% y
})[3]
```
---

## gpuR: Linear models
```{r}
# using gpuR
system.time({
  X <- gpuMatrix(runif(n*p), n, p)
  beta <- gpuMatrix(rnorm(p), p, 1)
  y <- X %*% beta + gpuMatrix(rnorm(n), n, 1)
  solve(t(X) %*% X) %*% t(X) %*% y
})[3]
```
---

## gpuR: SVD
```{r}
# using base R
n = 2000
p = 750
A <- matrix(runif(n*p), n, p)
gpuA <- gpuMatrix(A)

system.time({svd(A)})[3]

# using gpuR
system.time({svd(A)})[3]
```
---

## gpuR: Multivariate distance matrix regression
- Similar to regression, but using distance matrix to compute sum of squared distances<sup>1</sup>
- Calculation of pairwise distance between subjects is one of the most expensive steps
```{r, warning=FALSE}
# using base R
n = 1000; p = 1000; q = 10
system.time({
  X <- matrix(rnorm(n*q), n, q)
  Y <- matrix(runif(n*p), n, p)
  D <- dist(Y)
  mdmr(X, D)
})[3]
```
.footnote[<sup>1</sup>[Zapala and Schork, 2012](https://doi.org/10.3389/fgene.2012.00190)]
---

## gpuR: Multivariate distance matrix regression
```{r, warning=FALSE}
# using gpuR
system.time({
  X <- gpuMatrix(rnorm(n*q), n, q)
  Y <- gpuMatrix(runif(n*p), n, p)
  D <- dist(Y)
  mdmr(as.matrix(X), as.matrix(D))
})[3]
```
---

## gpuR: Summary
- Works for both AMD and Nvidia GPUs
- Requires minimal modification to code
- Can speed up many large matrix operations
- Often outperformed by Nvidia CUDA packages[<sup>1</sup>](http://www.parallelr.com/r-hpac-benchmark-analysis-gpu/)
---

## OpenCL in R
- `OpenCL` package provides a way to import OpenCL code
- Again, need to specify device
```{r include=FALSE}
library(OpenCL)
```

```{r, warning=FALSE}
library(OpenCL) # uses OpenCL, works on Mac
devices <- oclDevices() # choose device
ctx = oclContext(device = devices[[3]], precision = "best")
# precision="best" chooses double if available
```
---

## OpenCL: Importing very simple code
```{r}
# compute log of each vector element
code = c("
__kernel void logv(
  __global numeric* output,
 const unsigned int n,
  __global numeric* input)
{
  size_t i = get_global_id(0);
  if(i < n)
      output[i] = log(input[i]);
}")
# only allows for vector output
k_log <- oclSimpleKernel(ctx, "logv", code)
log_ocl <- function(x) as.numeric(oclRun(k_log, length(x), 
                                         as.clBuffer(x, ctx)))
```
---

## OpenCL: Testing log function
```{r}
a <- runif(2e8, 1, 2)
system.time({log(a)})[3]
system.time({log_ocl(a)})[3]
```
---

## OpenCL: Memory limitations
```{r, eval=FALSE}
system.time({log_ocl(runif(3e8, 1, 2))})[3]
```

```
Error in oclRun(k_log, length(x), as.clBuffer(x, ctx)) :  
Kernel execution failed (oclError -4)
```

- Error code defined in header file[<sup>1</sup>](https://tersetalk.wordpress.com/2012/04/13/opencl-error-codes/) as
```
CL_MEM_OBJECT_ALLOCATION_FAILURE
```
---

## Conclusions
- GPU is definitely useful for more than just gaming
- Substantial speedups in simple matrix operations
    - Limited advantage in complex tasks (e.g. SVD)
    - Unable to handle larger objects
- Easy to import OpenCL kernels
--
- Much more software available for Nvidia GPUs[<sup>1</sup>](https://www.r-pkg.org/ctv/HighPerformanceComputing)
    - Also see [`cuRnet`](https://github.com/InfOmics/cuRnet) for graph traversal
    - [`gpuMagic`](https://bioconductor.org/packages/release/bioc/html/gpuMagic.html) for a GPU-enabled `sapply` function
    - [`H2O4GPU`](https://cran.r-project.org/web/packages/h2o4gpu/vignettes/getting_started.html) for some basic machine learning functionalities