---
title: "Comparison of Multiple FC Harmonization Methods: Updated"
author:
  - Andrew Chen
  - Advised by Haochang Shou and Taki Shinohara
output:
  xaringan::moon_reader:
    css: ["PennSIVE-theme.css"]
    nature:
      beforeInit: "macros.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
header_includes:
  - \usepackage{amsmath}
  - \usepackage{bm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(patchwork)
library(reshape2)
library(dplyr)
library(tableone)
library(FCharmony)
library(kableExtra)
options("kableExtra.html.bsTable" = T)

plot_mat <- function(cov, lims = c(-max(abs(cov_melt$value)), max(abs(cov_melt$value)))) {
  dimnames(cov) <- list(NULL, NULL)
  cov_melt <- melt(cov)
  ggplot(data = cov_melt, aes(x=Var1, y=Var2, fill=value)) + 
    geom_tile() + 
    scale_fill_gradientn(
      colours = c("red", "white", "blue"), 
      limits = lims) +
    labs(fill = "") + 
    theme_classic() +
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
}
```

```{css}
.medium{font-size: 85%}
.small{font-size: 65%}
```

### Overview of FC Harmonization Methods
.footnote[
<sup>1</sup>[Yu et al., 2018](https://doi.org/10.1002/hbm.24241)  
<sup>2</sup>[Zhao et al., 2019](https://doi.org/10.1093/biostatistics/kxz057)]
- Harmonization of off-diagonal elements
    - ComBat applied directly<sup>1</sup>
--
- Log-transformation approaches
    - Apply PCA to off-diagonal elements then use ComBat on the scores
    - ComBat applied to the elements directly
--
- Harmonization of common principal component (CPC) eigenvalues
    - Apply ComBat to eigenvalues or log of eigenvalues
    - Assume log-linear relationship with covariates and regress out
    - Estimate CPCs using covariate-assisted principal (CAP) components<sup>2</sup>
---

### General Methodological Considerations
- Positive semi-definite constraints
    - Direct manipulation of FC matrices do not guarantee
    - Modification of log-transformed FC does
    - CPC-based methods can guarantee
--
- Correlation or covariance
    - Most downstream analyses rely on correlation
    - Can start with covariance or correlation input
    - No methods can guarantee correlation output
--
- Common structure among FC matrices
    - Empirical results show that majority of structure is not shared (CPC and PVD both confirm)
    - How to handle the error terms?
---

## Evaluation Criteria
- Norm-based approaches
    - Can evaluate pairwise distances between groups using any metric
    - MDMR can test by comparing within-group and between-group distances
--
- Network measures
    - Network connectivity: sum of all edges
    - Global efficiency: average of shortest path lengths
    - Local efficiency of a node: global efficiency of all connected nodes
        For a subgraph, we simply take the sum of all included local efficiencies
--
- Covariance regression models (WIP)
    - Could look at mass univariate testing of off-diagonal elements
    - Can also test for effects of site and covariates by assuming models on CPCs, CAPs, or other shared structures
---

## FC Harmonization using CPCs
Let $\Sigma_{ij}$ be the covariance matrices within site $i$ for subject $j$. We assume that there exist common principal components (CPCs) across the functional connectivity matrices such that
$$\Sigma_{ij} = \Phi \Lambda_{ij}\Phi^T + E_{ij}$$
where $\Lambda_{ij}$ are $p\times p$ diagonal matrices with entries $\lambda_{ijk}$ and $E_{ij}$ are the error terms.

We then directly apply the ComBat model to the log of $q$ CPC eigenvalues such that
$$\log\lambda_{ijk} = x_{ij}^T \beta_k + \gamma_{ik} + \delta_{ik} e_{ijk}$$
for $k=1,2,\ldots,q$ where $q \leq p$.
---

## Estimation of CPCs
- For now, $q$ is fixed at some value, but there could be a data-driven way (e.g. percentage of norm captured by CPC approximation)
- Estimation of CPCs is achieved through stepwise CPC<sup>1</sup>
- Currently testing three different approaches for handling the error terms
    - Do nothing
    - Element-wise ComBat
    - Remove entirely

.footnote[
<sup>1</sup>[Trendafilov, 2010](https://doi.org/10.1016/j.csda.2010.03.010)  
]
---

## Log-CovBat
- Apply log transformation then vectorize subject-specific matrices then perform PCA to obtain $\boldsymbol{\Phi}_k$, which are just the eigenvectors arranged as $p \times p$ symmetric matrices
$$\log\Sigma_{ij} = \sum_{k = 1}^K \Lambda_{ijk} \boldsymbol{\Phi}_k + \mathbf{E}_{ij}$$
- Apply ComBat including covariates to the $\Lambda_{ijk}$, harmonizing across sites indexed by $i$
- Recover CovBat-adjusted FC matrices by taking the matrix exponential
- Guarantees positive definiteness
- We also tried simply applying ComBat to the entries of $\log\Sigma_{ij}$
---

## All Methods Considered
- Bolded methods guarantee positive definiteness
    - For others, we find the nearest PD matrix
- Tried all methods using covariance and correlation as inputs


- FC ComBat: ComBat on Fisher-transformed off-diagonal elements
- **Log-ComBat**: Log then ComBat on all elements
- **Log-CovBat**: Log then ComBat on PC scores
- CPC eigenvalues harmonized by taking the log and applying ComBat
    - Not handled
    - ComBat on all elements
    - **Removed entirely**
---

## BLSA/CARDIA Data Example
- For initial tests, we are using a subset of data from the Baltimore Longitudinal Study of Aging (BLSA) and the Coronary Artery Risk Development in Young Adults (CARDIA) study
- Starting with 50 subjects per site, and random subset of 50 ROIs from the 264 available ROIs

```{r, include=FALSE}
load("../../BLSA CARDIA Example/blsa_cardia_small.Rdata")
```

```{r, include=FALSE}
BLSA_CARDIA_demo$site <- droplevels(BLSA_CARDIA_demo$site)
demotab <- CreateTableOne(vars = c("age_at_scan", "sex"), data = BLSA_CARDIA_demo, strata = "site")
tabout <- print(demotab)
rownames(tabout) <- c("Number of Subjects", "Age (mean(SD))", "Male (%)")
```

```{r demo, results="asis"}
knitr::kable(tabout[,1:4], "html") %>% kable_styling()
```
---

## Examples: Covariance
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- BLSA_CARDIA
s <- 1
site <- droplevels(BLSA_CARDIA_demo$site)
subj <- c(which(site == levels(site)[1])[s],
          which(site == levels(site)[2])[s],
          which(site == levels(site)[3])[s],
          which(site == levels(site)[4])[s])

plot_mat(fc[,,subj[1]]) + labs(title = "BLSA") + 
  plot_mat(fc[,,subj[2]]) + labs(title = "CARDIA1") +
  plot_mat(fc[,,subj[3]]) + labs(title = "CARDIA3") +
  plot_mat(fc[,,subj[4]]) + labs(title = "CARDIA4")
```

---
## Examples: Correlation
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## FC ComBat Correlations
```{r}
load("../../BLSA CARDIA Example/corr_dats.Rdata")
```

```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- yu_out$dat.out
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## Log-ComBat Correlations
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- log_c_out$dat.out
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## Log-CovBat Correlations
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- log_out$dat.out
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## CPC+None Correlations
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- cpc_out$dat.out
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## CPC+ComBat Correlations
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- cpc_c_out$dat.out
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## CPC+Remove Correlations
```{r fig.width=8, fig.height=7, fig.show='hold', fig.align='center', out.width='80%', fig.retina=3}
fc <- cpc_r_out$dat.out
plot_mat(cov2cor(fc[,,subj[1]])) + labs(title = "BLSA") + 
  plot_mat(cov2cor(fc[,,subj[2]])) + labs(title = "CARDIA1") +
  plot_mat(cov2cor(fc[,,subj[3]])) + labs(title = "CARDIA3") +
  plot_mat(cov2cor(fc[,,subj[4]])) + labs(title = "CARDIA4")
```
---

## MDMR Results: Covariance Inputs
```{r}
load("all_cov_harmony_out.Rdata")
```
- FC ComBat excluded since it only operates on correlation matrices
- MDMR implemented via 10,000 permutations of the independent variables
- Results reported for both covariance and correlation outputs

.center[Results for Frobenius Distance]
.small[
```{r , results="asis"}
ccolors <- colorRampPalette(c("#011F5B", "#000000", "#990000"))(256)
color_kable <- function(x) {
  colored <- as.data.frame(x) %>% 
    mutate_if(is.numeric, function(y) {
      cell_spec(y, "html", 
                color = ccolors[round(scales::rescale(y, to = c(1, 256), from = 0:1))])}
    )
  rownames(colored) <- rownames(x)
  knitr::kable(colored, "html", escape = FALSE, row.names = TRUE)
}

colnames(all_reg) <- c("All", "Age", "Sex", "Site", 
                         "All.Corr", "Age.Corr", "Sex.Corr", "Site.Corr")
colnames(all_regL) <- c("All", "Age", "Sex", "Site", 
                         "All.Corr", "Age.Corr", "Sex.Corr", "Site.Corr")
color_kable(all_reg)
```
]
---

## MDMR Results: Covariance Inputs
.center[Results for Log-Euclidean Distance]
.small[
```{r , results="asis"}
color_kable(all_regL)
```
]
---

## MDMR Results: Correlation Inputs
```{r}
load("all_corr_harmony_out.Rdata")
```
- Results reported for correlation inputs and outputs
- For log-Euclidean distance, had to find the nearest PD matrix to FC ComBat output

.center[Results for Frobenius Distance]
```{r , results="asis"}
colnames(all_reg) <- c("All", "Age", "Sex", "Site")
colnames(all_regL) <- c("All", "Age", "Sex", "Site")
color_kable(all_reg)
```
---

## MDMR Results: Correlation Inputs
.center[Results for Log-Euclidean Distance]
```{r , results="asis"}
color_kable(all_regL)
```
---

## MDMR Results: Covariance Inputs
.center[Results for Log-Euclidean Distance]
.small[
```{r , results="asis"}
color_kable(all_regL)
```
]
---

## Network Results: Covariance Inputs
```{r}
load("all_cov_harmony_out.Rdata")
```
- Results reported for covariance inputs and correlation outputs

.center[Results for Covariate Association]
.small[
```{r , results="asis"}
color_kable(round(all_yu$net.results,2)[,-1])
```
]
---

## Network Results: Covariance Inputs
.center[Results for Site Association]
.small[
```{r , results="asis"}
color_kable(round(all_yu$net.results.site,2))
```
]
---

## Network Results: Covariance Inputs
```{r}
load("all_corr_harmony_out.Rdata")
```
- Results reported for correlation inputs and outputs

.center[Results for Covariate Association]
.small[
```{r , results="asis"}
color_kable(round(all_yu$net.results,2)[,-1])
```
]
---

## Network Results: Covariance Inputs
.center[Results for Site Association]
.small[
```{r , results="asis"}
color_kable(round(all_yu$net.results.site,2))
```
]
---

## Summary
- Results depend heavily on the choice of metric
    - Promising results for CPC+ComBat and log-ComBat which show strong harmonization performance in both metrics
- Need to implement and test CPC-based regression methods
- Adapting CAP code for application in harmonization
- Further testing of network metrics
- Working on an R package to implement all methods and tests

```{r, eval=FALSE}
pagedown::chrome_print() # print as pdf
```

